---
title: "ManyPrimates plots"
author: "Manuel Bohn"
date: "4 7 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(tidyverse)
library(ggthemes)
library(langcog)

# langcog package is installed by running
# install.packages("devtools")
# devtools::install_github("langcog/langcog")

```

# Intro

These are some preliminary visualizations of the data.

```{r loading data, include = FALSE}
mp_data <- read.csv(file = "../data/merged_data/01_manyprimates_pilot_merged_data.csv")
```

# Overview by species

First an overview of the data, plotted by species and condition. Small, transparent dots represent aggregated data for each individual by condition. Solid dots is the group mean for that condition. Error bars are 95% confidence intervals.

```{r overview by species}
plot_individual <- mp_data %>%
  mutate(phylo = case_when(species == "chimpanzee" ~ 9,
                           species == "ring_tailed_lemur" ~ 1,
                           species == "orangutan"~ 12,
                           species == "bonobo"~ 10,
                           species == "gorilla"~ 11,
                           species == "black_and_white_ruffed_lemur"~ 2,
                           species == "brown_capuchin_monkey" ~ 3,
                           species == "black_faced_spider_monkey" ~ 4,
                           species == "long_tailed_macaque" ~ 6,
                           species == "barbary_macaque" ~ 7,
                           species == "rhesus_macaque" ~ 8,
                           species == "squirrel_monkey"~ 5),
         species = reorder(species,phylo))%>%
  group_by(species,condition,subject_site) %>%
  summarise(correct = mean(correct),
             nr_trials = max(trial))


plot_group <- plot_individual %>%
   multi_boot_standard(col = "correct")


ggplot() +
  geom_jitter(data = plot_individual, aes(x = condition, y = correct, col = condition, size = nr_trials), width = .3,height = .015, alpha = .2)+
  geom_pointrange(data = plot_group, aes(x = condition, y = mean, col = condition,ymin = ci_lower, ymax = ci_upper),size = 1.5)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="proportion correct")+
  facet_grid(species ~ . ,switch = "y") +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  theme(axis.text.y=element_blank(),axis.ticks.y=element_blank(),strip.text.y = element_text(angle = 180))+
  scale_colour_solarized(name="delay")+
  coord_flip()
```

# Plots by site

Here we select the species for which we have data from multiple sites. This is a very preliminary way of checking whether there is a lot of variation between sites. Plotting conventions are the same as above.

First we check for which species we have data from more than one site:
```{r plots by site}
# plotting data by site for species for which we have data from more than one lab

# check for which species this applies

mp_data%>%
  group_by(species)%>%
  summarise(sites = length(unique(site)))%>%
  knitr::kable()
```

## Chimpanzees

```{r plots by site chimps}
# starting with chimps

chimp_plot_individual <- mp_data %>%
  filter(species == "chimpanzee")%>%
  group_by(site,condition,subject_site) %>%
  summarise(correct = mean(correct))

chimp_plot_group <- chimp_plot_individual %>%
   multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = chimp_plot_individual, aes(x = condition, y = correct, col = condition),width = .3,height = .015, alpha = .5)+
  geom_pointrange(data = chimp_plot_group, aes(x = condition, y = mean, col = condition,ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="proportion correct")+
  facet_grid(site ~ . ,switch = "y") +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  theme(axis.text.y=element_blank(),axis.ticks.y=element_blank(),strip.text.y = element_text(angle = 180))+
  scale_colour_solarized(name="delay")+
  coord_flip()

```

## Ring-tailed lemurs

```{r plots by site ring-tailed-lemurs}
rtlemur_plot_individual <- mp_data %>%
  filter(species == "ring_tailed_lemur")%>%
  group_by(site,condition,subject_site) %>%
  summarise(correct = mean(correct))

rtlemur_plot_group <- rtlemur_plot_individual %>%
   multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = rtlemur_plot_individual, aes(x = condition, y = correct, col = condition),width = .3,height = .015, alpha = .5)+
  geom_pointrange(data = rtlemur_plot_group, aes(x = condition, y = mean, col = condition,ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="proportion correct")+
  facet_grid(site ~ . ,switch = "y") +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  theme(axis.text.y=element_blank(),axis.ticks.y=element_blank(),strip.text.y = element_text(angle = 180))+
  scale_colour_solarized(name="delay")+
  coord_flip()

```


## Brown Capuchins

```{r plots by site capuchins}
cap_plot_individual <- mp_data %>%
  filter(species == "brown_capuchin_monkey")%>%
  group_by(site,condition,subject_site) %>%
  summarise(correct = mean(correct))

cap_plot_group <- cap_plot_individual %>%
   multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = cap_plot_individual, aes(x = condition, y = correct, col = condition),width = .3,height = .015, alpha = .5)+
  geom_pointrange(data = cap_plot_group, aes(x = condition, y = mean, col = condition,ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="proportion correct")+
  facet_grid(site ~ . ,switch = "y") +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  theme(axis.text.y=element_blank(),axis.ticks.y=element_blank(),strip.text.y = element_text(angle = 180))+
  scale_colour_solarized(name="delay")+
  coord_flip()

```

## Bonobos

```{r plots by site bonobo}
bon_plot_individual <- mp_data %>%
  filter(species == "bonobo")%>%
  group_by(site,condition,subject_site) %>%
  summarise(correct = mean(correct))

bon_plot_group <- bon_plot_individual %>%
   multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = bon_plot_individual, aes(x = condition, y = correct, col = condition),width = .3,height = .015, alpha = .5)+
  geom_pointrange(data = bon_plot_group, aes(x = condition, y = mean, col = condition,ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="proportion correct")+
  facet_grid(site ~ . ,switch = "y") +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  theme(axis.text.y=element_blank(),axis.ticks.y=element_blank(),strip.text.y = element_text(angle = 180))+
  scale_colour_solarized(name="delay")+
  coord_flip()

```


# Task-experience

Here we split each species by task experience. Check if we have species with sufficient memebers having different levels of task experience.

```{r plots by task experience}
# check for which species this applies

mp_data%>%
  group_by(species, task_experience)%>%
  summarise(n = length(unique(subject_site)))%>%
  knitr::kable()
```


So far, this only applies to chimps and bonobos. However, task experience co-varies with site.

## Chimpanzees

```{r plots by task experience chimps}
# starting with chimps

chimp_task_plot_individual <- mp_data %>%
  filter(species == "chimpanzee")%>%
  group_by(site, task_experience,condition,subject_site) %>%
  summarise(correct = mean(correct))

chimp_task_plot_group <- mp_data %>%
  filter(species == "chimpanzee")%>%
  group_by(task_experience,condition,subject_site) %>%
  summarise(correct = mean(correct))%>%
   multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = chimp_task_plot_individual, aes(x = condition, y = correct, col = site),width = .3,height = .015, alpha = .7)+
  geom_pointrange(data = chimp_task_plot_group, aes(x = condition, y = mean ,ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="proportion correct")+
  facet_grid(task_experience ~ . ,switch = "y") +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  theme(axis.text.y=element_blank(),axis.ticks.y=element_blank(),strip.text.y = element_text(angle = 180))+
  scale_colour_hue(name="site")+
  coord_flip()

```

## Bonobos

```{r plots by task experience chimps}
bonobo_task_plot_individual <- mp_data %>%
  filter(species == "bonobo")%>%
  group_by(site, task_experience,condition,subject_site) %>%
  summarise(correct = mean(correct))

bonobo_task_plot_group <- mp_data %>%
  filter(species == "bonobo")%>%
  group_by(task_experience,condition,subject_site) %>%
  summarise(correct = mean(correct))%>%
   multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = bonobo_task_plot_individual, aes(x = condition, y = correct, col = site),width = .3,height = .015, alpha = .7)+
  geom_pointrange(data = bonobo_task_plot_group, aes(x = condition, y = mean ,ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="proportion correct")+
  facet_grid(task_experience ~ . ,switch = "y") +
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  theme(axis.text.y=element_blank(),axis.ticks.y=element_blank(),strip.text.y = element_text(angle = 180))+
  scale_colour_hue(name="site")+
  coord_flip()

```


# Age

Here we plot age against correct choice separate for each condition and species. Regression line is smoothed conditional mean.

```{r plotting age}
plot_age <- mp_data %>%
   group_by(subject_site,norm_age,species,condition) %>%
   summarise(correct = mean(correct))


ggplot(plot_age, 
       aes(x = norm_age, y = correct))+
  geom_point(aes(colour=species))+
  geom_smooth(method = "loess", se = F , level = 0.95)+
  guides(fill=FALSE)+
  geom_hline(yintercept = 1/3, lty=2)+
  #facet_grid( condition ~ species )+
  facet_grid( condition ~ . )+
  theme_few()
```

# Cup distance (sig. effect in preliminary model)

```{r plotting cup distance}
plot_cup <- mp_data %>%
   group_by(subject_site,cup_distance,species,condition) %>%
   summarise(correct = mean(correct))


ggplot(plot_cup, 
       aes(x = cup_distance, y = correct))+
  geom_jitter(aes(colour=species))+
  geom_smooth(method = "loess", se = F , level = 0.95)+
  guides(fill=FALSE)+
  geom_hline(yintercept = 1/3, lty=2)+
  theme_few()
```
