---
title: "ManyPrimates inference"
author: "Manuel Bohn"
date: "10 Oct 2018"
output:
  html_notebook:
    code_folding: hide
    css: style.css
    theme: paper
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(lsr)
```

# Intro

Preparing code for inferential statistics according to pre-registration analysis. 

```{r loading data}
mp_data <- read_csv("../data/merged_data/01_manyprimates_pilot_merged_data_v2.csv", 
                    col_types = list(age = 'd', cup_distance = 'd'))
```

## T-tests against chance per species and delay type

Quote from Registration:

> T-tests: Proportion of correct choices different from chance level (0.33) per species and delay type.
> 
> For this analysis, performance is aggregated per individual.

For species with only 1 subject, t-tests cannot be computed. They are therefore exluced.

```{r comparisons to chance}
mp_data %>%
  # no t test possible because ...
  filter(species != "black_faced_spider_monkey", # only 1 data point so far
         !(species == 'orangutan' & delay == 'short')) %>% # data constant at M=1 (i.e., SD=0)
  mutate(delay = fct_rev(delay)) %>%
  group_by(species, delay, subject_site) %>%
  summarise(correct = mean(correct)) %>%
  summarise(n = n(),
         mean = mean(correct) %>% round(2),
         # df = t.test(correct, mu = 1/3)$parameter, # redundant info in combination with n
         t = t.test(correct, mu = 1/3)$statistic %>% round(2),
         p = t.test(correct, mu = 1/3)$p.value %>% round(3),
         d = cohensD(correct, mu = 1/3) %>% round(2)) %>%
  mutate(p.visual = case_when(p<.001 ~ '***', p<.01 ~ '**', p<.05 ~ '*', p<.01 ~ '.', T ~ '')) %>%
  select(species:mean, d, everything()) #%>%
  # knitr::kable(digits = 2)
```

