---
title: "ManyPrimates inference"
author: "Manuel Bohn"
date: "8 7 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(tidyverse)

```

# Intro

Preparing code for inferential statistics according to pre-registration analysis. 

```{r loading data, include = FALSE}
mp_data <- read_csv(file = "../data/merged_data/01_manyprimates_pilot_merged_data.csv")
```

## T-tests against chance per species and delay type

Quote from Registration:

> T-tests: Proportion of correct choices different from chance level (0.33) per species and delay type.
For this analysis, performance is aggregated per individual.

For species with only 1 subject, t-tests cannot be computed. They are therefore exluced.

```{r comparisons to chance}
mp_data %>%  
  filter(species != "black_faced_spider_monkey")%>% # only one data point so far, therefor no t test possible
  group_by(species ,condition, subject_site) %>%
  summarise(correct = mean(correct)) %>%
  summarize(correct = list(correct)) %>%
  group_by(species ,condition) %>%
  mutate(n = length(unlist(correct)),
         mean = mean(unlist(correct)),
  #       df = t.test(unlist(correct), mu = 1/3)$parameter, # redundant info in combination with n
         t = t.test(unlist(correct), mu = 1/3)$statistic,
         p = t.test(unlist(correct), mu = 1/3)$p.value) %>%
  select(species ,condition,n,mean,t,p)%>%
  knitr::kable(digits = 2)
```

