---
title: "ManyPrimates pilot mixed modeling"
author: "Drew Altschul"
date: "13 July 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)
library(readr)
library(tidyverse)
library(sjPlot)

```

## Data import

Preparing code for pre-registered mixed modeling. 

```{r loading data, include = FALSE}
setwd('M:/SourceTree/ManyPrimates_pilot/analysis/') # only needed for DMA's setup - should be deleted eventually
mp_data <- read.csv(file = "../data/merged_data/01_manyprimates_pilot_merged_data.csv")
```



## Mixed modeling with all relevant varaibles predicting accuracy

From the preregistration, the mixed model was specified thusly:

> correct ~ delay * age + 
task_experience + cup_distance + board_size + trial +
         (1 + delay + trial | site/subject/block/hiding_location ) + 
        (1 + task_experience + cup_distance + board_size + trial + delay | species)


In the dataframe, 
condition = delay,
subject_site = subject,
and norm_age should be used for age.

Model as pre-registered has too many random effects

> Error: number of observations (=5807) < number of random effects (=10108) for term (1 + condition + trial | hiding_location:(block:(subject_site:site))); the random-effects parameters are probably unidentifiable

Pruning random effects in the following order (from preregistration): 
> * Remove correlations between random effects
  * Remove random slopes (in the following order)
  + Site
  + Species
  + Hiding_location
  + Block
  + Subject

Model only converges once we take out `hiding_location`. After doing so, the other random effects (correlation, site, species) can be put back in.

The model below converges. Model output is saved in `06_mp_model.rds`

``` {r mixed modeling, error=TRUE}

# centering variables for modeling
# filtering spider monkey. Only one data point so far, therefore this is not worth including to explode the number of random effects

model.data <- mp_data %>%
  filter(species != "black_faced_spider_monkey")%>%
  mutate(cup_distance = scale(cup_distance, scale = T, center = T),
         board_size = scale(board_size, scale = T, center = T),
         trial = scale(trial, scale = T, center = T),
         hiding_location = as.factor(hiding_location),
         condition = relevel(condition, ref = "short"))

## Model takes a while to run. Run next line to load model output from previous run with structure below.

mm.1 <- readRDS("06_mp_model.rds")

## some diagnostics

## examining Cholesky decomposition
theta <- getME(mm.1,"theta")
diag.element <- getME(mm.1,"lower")==0
any(theta[diag.element]<1e-5)


## model summary

summary(mm.1)

## code to run the model
mm.1 <- glmer(correct ~ condition * norm_age + 
               task_experience + cup_distance + board_size + trial +
               (1 + condition + trial || site/subject_site/block ) +         
               (1 + task_experience + cup_distance + board_size + trial + condition | species)
             , data = model.data
             , family = binomial
             , control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)

## CIs
confint(mm.1, method='profile')


## save model output
saveRDS(mm.1, "06_mp_model.rds")

### Model plotting

plot_model(mm.1, type='est') 
plot_model(mm.1, type='re') 
## sjp.lmer will be deprecated eventually, and plot_model can be used to precisely obtain which plots of the GLMM one wishes to examine.
sjp.lmer(mm.1)
## In the meantime, sjp.lmer will plot the random effects by default.


```

```{r}
# reduced model

## Model takes a while to run. Run next line to load model output from previous run with structure below.

mm.2 <- readRDS("06_2_mp_model2.rds")

mm.2.ci<- readRDS("06_2_mp_model2_ci.rds")

# Pruning the model

## remove block from random effects as the estimates in the previous models were essentially 0
## put correlation between random effects within site/species back in 

  #correct ~ condition * norm_age + 
  #             task_experience + cup_distance + board_size + trial +
  #             (1 + condition + trial | site/subject_site ) +         
  #             (1 + task_experience + cup_distance + board_size + trial + condition | species)

## check for colinearity in the previous model.

library (car)

col.mm1 <- glm(correct ~ condition + norm_age + 
                 task_experience + cup_distance + board_size + trial
               , data = model.data
               , family = binomial)

vif(col.mm1)

### board size and cup distance show high colinearity. 
### removing board size as it is highly correlated with cup distance. Cup distance seems to be of more immediate relevance.

  #correct ~ condition * norm_age + 
  #             task_experience + cup_distance + trial +
  #             (1 + condition + trial | site/subject_site ) +         
  #             (1 + task_experience + cup_distance + trial + condition | species)

## check levels of random effects
### the following function checks how many different levels there are within  each random effect

source("../../diagnostic_fcns.r") 

Overview=fe.re.tab("correct ~ condition + task_experience + cup_distance + board_size + trial", "species",data = model.data)

Overview$summary

#### this suggests that random slopes for task experience does not make much sense as most species have only 1 level. Same is true for cup distance size. Therefore they are removed

  #correct ~ condition * norm_age + 
  #             task_experience + cup_distance + board_size + trial +
  #             (1 + condition + trial | site/subject_site ) +         
  #             (1 + trial + condition | species)

## run model: trial random slopes within species also yields random effects equal to 0. Therefore it is removed 

# reduced model 

mm.3 <- glmer(correct ~ condition * norm_age + species +
                task_experience + cup_distance +  trial +
                (1 + trial + condition | site / subject_site ) +         
                (1 + condition | species)
              , data = model.data
              , family = binomial
              , control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)

summary(mm.2)

# CIs
# does not work ?
confint(mm.2, method='profile')

# Bootsrap function by Roger Mundry @ MPI EVA

source("../../boot_glmm.r")

mm.2.ci = boot.glmm.pred(model.res=mm.2, excl.warnings=F, nboots=1000, para=F, resol=100, level=0.95, use=NULL, circ.var.name=NULL, circ.var=NULL, use.u=F,n.cores=c("all-1", "all"), save.path=NULL)


# save model output
saveRDS(mm.2, "06_2_mp_model2.rds")
saveRDS(mm.2.ci, "06_3_mp_model3_drop1.rds")


### Model plotting

plot_model(mm.2, type='est', title = "Fixed Effects")+
  geom_hline(yintercept = 1, lty=2)+
  ylim(.05,3)+
  theme_few()

plot_model(mm.2, type='re') 


```


```{r}
a <-tidy(mm.2)%>%
  filter(term == "conditionlong" |term == "conditionmedium" | term == "cup_distance")%>%
  mutate(estimate.m2 = round(estimate,2))%>%
  select(term,estimate.m2)
  

b <-tidy(mm.1)%>%
  filter(term == "conditionlong" |term == "conditionmedium" | term == "cup_distance")%>%
  mutate(estimate.m1 = round(estimate,2))%>%
  select(term,estimate.m1)

b %>%
  mutate(diff = estimate.m1 - a$estimate.m2)




```
```{r}
mm.3 <- glmer(correct ~ condition * norm_age +
                task_experience + cup_distance +  trial +
       
                (1 + condition | species)
              , data = model.data
              , family = binomial
              , control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
)


plot_model(mm.3, type='re', sort.est = T) 

```

