---
title: "Bayesian analysis MPP"
output: html_document
---

```{r}
library(tidyverse)
library(brms)
library(emmeans)
library(car)
```
```{r loading data}
mp_data <- read.csv("../data/merged_data/01_manyprimates_pilot_merged_data_v2.csv")
```

```{r}
model.data <- mp_data %>%
  filter(species != "black_faced_spider_monkey") %>%
  mutate_at(vars(cup_distance, board_size, trial), funs(scale(.)[, 1])) %>%
  mutate(hiding_location = factor(hiding_location),
         delay = fct_relevel(delay, "short"))
```

We will include random slopes for those predictors that have vary in more than 50% of the respective intercept term. Here we chech for `cup_distance`, `board_size` and `task_experience` within species. 

```{r}
mp_data %>%
  group_by(species)%>%
  summarise(levels_cup = length(unique(cup_distance)),
            levels_board = length(unique(board_size)),
            levles_task = length(unique(task_experience)))%>%
  summarise(cup = sum(levels_cup > 1) /length(levels_cup),
            board = sum(levels_board > 1) /length(levels_board),
            task = sum(levles_task > 1) /length(levles_task))
  
```

They vary less and so we will not include those random slopes within species. 

Next we check collinearity between predictors - tehre is no bayesian version for that so we just run a `glm`
```{r}
vif_model <- glm(correct ~ delay + norm_age +
                 task_experience + cup_distance + board_size + trial
               , data = model.data
               , family = binomial)

vif(vif_model)
```

looks fine, so we include all predictors. 

```{r}
get_prior(
  correct ~ delay + 
    task_experience + norm_age + cup_distance + board_size + trial +
    (1 + delay + trial | site/subject_site ) +
    (1 + trial + delay | species),
  family = bernoulli(),
  data = model.data
)
```

Set priors for the model

```{r}
prior <- c(
    prior(normal(0, 2), class = Intercept),
    prior(normal(0, 2), class = b),
    prior(lognormal(0, 1), class = sd, group = site),
    prior(lognormal(0, 1), class = sd, group = `site:subject_site`),
    prior(lognormal(0, 2), class = sd, group = species),
    prior(lkj(2), class = cor)
    )
```

check if the priors made it into the model

```{r}
make_stancode(correct ~ delay + 
    task_experience + norm_age + cup_distance + board_size + trial +
    (1 + delay + trial | site/subject_site ) +
    (1 + trial + delay | species),
  family = bernoulli(),
  data = model.data,
  prior = prior)
```

run the model

```{r}
bm <- brm(
  correct ~ delay + 
    task_experience + norm_age + cup_distance + board_size + trial +
    (1 + trial + delay | site/subject_site ) +
    (1 + trial + delay | species),
  family = bernoulli(),
  data = model.data,
  prior = prior,
  iter = 5000,
  warmup = 2000,
  cores = 4,
  chains = 4,
  save_all_pars = T, 
  sample_prior = TRUE
)

```

 Links: mu = logit 
Formula: correct ~ delay + task_experience + norm_age + cup_distance + board_size + trial + (1 + trial + delay | site/subject_site) + (1 + trial + delay | species) 
   Data: model.data (Number of observations: 5850) 
Samples: 4 chains, each with iter = 5000; warmup = 2000; thin = 1;
         total post-warmup samples = 12000

Group-Level Effects: 
~site (Number of levels: 12) 
                           Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)                  0.75      0.24     0.37     1.31 1.00     3825     5731
sd(trial)                      0.09      0.07     0.00     0.24 1.00     4053     5385
sd(delaylong)                  0.20      0.16     0.01     0.61 1.00     3320     4899
sd(delaymedium)                0.32      0.18     0.02     0.74 1.00     2910     2495
cor(Intercept,trial)           0.16      0.37    -0.61     0.78 1.00    11700     8384
cor(Intercept,delaylong)      -0.15      0.38    -0.81     0.61 1.00     6484     8338
cor(trial,delaylong)           0.02      0.38    -0.69     0.71 1.00    10347     9242
cor(Intercept,delaymedium)    -0.31      0.33    -0.84     0.42 1.00     8146     7976
cor(trial,delaymedium)        -0.03      0.37    -0.70     0.67 1.00     8277     9065
cor(delaylong,delaymedium)     0.17      0.38    -0.61     0.82 1.00     5582     8536

~site:subject_site (Number of levels: 175) 
                           Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)                  0.62      0.09     0.47     0.83 1.00     1893     3410
sd(trial)                      0.07      0.05     0.00     0.18 1.00     3471     5007
sd(delaylong)                  0.30      0.16     0.02     0.63 1.00      992     2242
sd(delaymedium)                0.20      0.14     0.01     0.52 1.00     1091     1814
cor(Intercept,trial)          -0.02      0.34    -0.67     0.66 1.00    13230     8838
cor(Intercept,delaylong)      -0.53      0.28    -0.89     0.20 1.00     3018     4047
cor(trial,delaylong)           0.04      0.37    -0.67     0.72 1.00     4076     6828
cor(Intercept,delaymedium)    -0.21      0.36    -0.79     0.56 1.00     2617     6217
cor(trial,delaymedium)         0.03      0.38    -0.69     0.72 1.00     3991     7439
cor(delaylong,delaymedium)     0.30      0.40    -0.57     0.88 1.00     1555     3776

~species (Number of levels: 11) 
                           Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)                  0.60      0.22     0.27     1.14 1.00     4089     5185
sd(trial)                      0.06      0.05     0.00     0.20 1.00     4825     4610
sd(delaylong)                  0.32      0.16     0.05     0.68 1.00     3503     2789
sd(delaymedium)                0.24      0.15     0.01     0.59 1.00     3913     3565
cor(Intercept,trial)          -0.02      0.37    -0.71     0.67 1.00    13397     9400
cor(Intercept,delaylong)      -0.41      0.32    -0.88     0.32 1.00     6195     7630
cor(trial,delaylong)           0.10      0.37    -0.63     0.75 1.00     7293     9108
cor(Intercept,delaymedium)    -0.31      0.35    -0.86     0.45 1.00     8503     8491
cor(trial,delaymedium)         0.03      0.37    -0.69     0.71 1.00    10872     9985
cor(delaylong,delaymedium)     0.32      0.35    -0.47     0.87 1.00     7156     7202

Population-Level Effects: 
                   Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept              1.27      0.40     0.44     2.04 1.00     4042     5468
delaylong             -1.21      0.19    -1.59    -0.84 1.00     5033     6030
delaymedium           -0.93      0.19    -1.30    -0.55 1.00     5567     7054
task_experienceyes     0.36      0.31    -0.23     1.01 1.00     4498     5695
norm_age               0.04      0.07    -0.09     0.18 1.00     6272     7893
cup_distance           0.49      0.24     0.02     0.99 1.00     4521     5368
board_size             0.46      0.21     0.04     0.88 1.00     4456     5902
trial                  0.01      0.06    -0.11     0.12 1.00     6485     7010

Samples were drawn using sampling(NUTS). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).


compare the different delay conditions

```{r}
emmeans(bm, pairwise ~ delay)$contrasts%>%
  as_tibble()%>%
  ggplot(., aes(x = contrast, y = estimate))+
  geom_hline(yintercept = 0, lty = 2, alpha = .5)+
  geom_pointrange(aes(ymin = lower.HPD, ymax = upper.HPD))+
  coord_flip()+
  ylim(-2,2)+
  theme_minimal()
```

```{r}
x<-seq(-10,10,by=0.1)
plot(x, dnorm(x, mean = 0, sd = 1),type="l", ylim=c(0, 0.8))
#lines(x, dcauchy(x, location = 0, scale = 1),col="red")
#lines(x, dstudent_t(x, 3, 0, 2.5),col="green")
lines(x, dlnorm(x, 0,1),col="firebrick")
```



